\documentclass[12pt,a4paper]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{longtable}

\titleformat{\section}{\large\bfseries\centering}{\thesection.}{0.5em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{0.5em}{}

\begin{document}

\begin{center}
{\Large\bfseries PATENT APPLICATION}\\[0.5em]
{\large UNITED KINGDOM INTELLECTUAL PROPERTY OFFICE}\\[1.5em]
{\Large\bfseries Adaptive Multi-Spectrum Anomaly Detection, Risk-Aware\\
Transaction Processing, and Cross-Domain Stability Control System}\\[1.5em]
{\large Inventor: Odeyemi Olusegun Israel}\\[0.5em]
{\large Date of Filing: \today}\\[0.5em]
{\large Applicant: Odeyemi Olusegun Israel}\\[0.3em]
{\large Nationality: Nigerian}\\[0.3em]
{\large Address: [INSERT ADDRESS]}
\end{center}

\vspace{1em}
\hrule
\vspace{1em}

%% ============================================================
%% ABSTRACT
%% ============================================================
\section*{ABSTRACT}

A unified system and method for detecting anomalies, scoring risk, enforcing
compliance, and controlling stability across heterogeneous domains.  The
invention comprises: (1)~a Blind-Spot Decomposition Theory (BSDT) that
decomposes classification failures into four independent diagnostic channels and
produces a Missed Fraud Likelihood Score enabling post-hoc classifier correction
without retraining; (2)~a Universal Deviation Principle (UDP) employing
multi-spectrum operator families with provable detection guarantees and
semidefinite-programming-based optimal fusion; (3)~an N-body variational scoring
engine (GravityEngine) with closed-form energy potentials for anomaly
separation; (4)~a boundary-centred dimension magnifier that nonlinearly
amplifies discriminative features while silencing non-discriminative ones;
(5)~a risk-aware blockchain transaction processing protocol with tiered escrow,
ML-oracle risk routing, zero-knowledge compliance verification, and cross-chain
risk propagation; (6)~a Lyapunov-based adaptive friction stability framework
for systemic risk detection and optimal macroprudential policy; (7)~a
state-dependent viscosity method transferring the BSDT functional to
computational fluid dynamics; and (8)~a UI integrity verification service
preventing interface manipulation attacks.  The system operates across financial
fraud detection, blockchain compliance, macroeconomic stability, and
mathematical physics, unified by a common blind-spot energy functional.

\newpage

%% ============================================================
%% DESCRIPTION
%% ============================================================
\section{FIELD OF THE INVENTION}

This invention relates to computational methods for anomaly detection, risk
scoring, transaction compliance enforcement, systemic stability control, and
adaptive viscosity in fluid dynamics.  More particularly, it relates to a
unified mathematical framework that decomposes classifier blind spots into
independent diagnostic channels, and applies the resulting energy functional
across multiple domains including financial fraud detection, blockchain
transaction processing, macroeconomic stability monitoring, and Navier--Stokes
simulation.

\section{BACKGROUND OF THE INVENTION}

Machine learning classifiers deployed for fraud detection exhibit systematic
failure modes that are poorly understood.  Existing approaches treat missed
detections as residual noise rather than structurally decomposable phenomena.
This leads to costly retraining cycles, undetected fraud, and inability to
transfer detection insights across domains.

In blockchain systems, transaction compliance requires reconciling
anti-money-laundering (AML) requirements with privacy guarantees and
decentralisation constraints.  No existing system integrates ML risk scoring,
zero-knowledge cryptographic proofs, cross-chain policy synchronisation, and
decentralised dispute resolution into a unified protocol.

In macroeconomic regulation, countercyclical capital buffers are set using
lagging indicators without formal stability guarantees.  In computational fluid
dynamics, artificial viscosity methods lack principled diagnostic criteria for
where and when to activate.

The present invention addresses all of these problems through a common
mathematical framework.

\section{SUMMARY OF THE INVENTION}

The invention provides a computer-implemented system and method comprising
eight interrelated subsystems, unified by a common blind-spot energy functional
$E_{\mathrm{BS}}$.  The subsystems are:

\begin{enumerate}[label=(\alph*)]
\item \textbf{Blind-Spot Decomposition Theory (BSDT):} A four-channel
  diagnostic framework decomposing classification failures into Camouflage~($C$),
  Feature Gap~($G$), Activity Anomaly~($A$), and Temporal Novelty~($T$)
  components, with a composite Missed Fraud Likelihood Score (MFLS) enabling
  post-hoc correction.

\item \textbf{Universal Deviation Principle (UDP):} A multi-spectrum anomaly
  detection framework using families of deviation-separating operators with
  provable full-rank detection guarantees, SDP-optimal fusion, and submodular
  greedy operator selection.

\item \textbf{GravityEngine:} An N-body variational anomaly scoring algorithm
  with closed-form energy potentials, force--energy consistency, and
  backtracking line-search convergence.

\item \textbf{Dimension Magnifier:} A boundary-centred nonlinear per-dimension
  transform using a saturating function whose steepness is proportional to
  per-dimension discriminative score.

\item \textbf{Risk-Aware Transaction Protocol (AMTTP):} A blockchain protocol
  integrating ML-oracle risk scoring with tiered escrow
  (Approve/Review/Escrow/Block), hash time-locked contracts, policy-engine
  enforcement, and circuit breakers.

\item \textbf{Zero-Knowledge Compliance (zkNAF):} Privacy-preserving
  compliance verification using Groth16 ZK-SNARK proofs for sanctions
  non-membership, risk-range attestation, and KYC verification without data
  disclosure.

\item \textbf{Adaptive Friction Stability Framework:} A Lyapunov-based
  systemic risk detection and optimal macroprudential policy derivation system,
  with spectral phase-transition criteria, closed-form optimal damping, and
  online regret-bounded policy adjustment.

\item \textbf{Cross-Domain Transfer:} Application of the BSDT energy
  functional as state-dependent viscosity in Navier--Stokes simulations, and
  as the risk metric in cross-chain propagation, UI integrity verification,
  and multi-service compliance orchestration.
\end{enumerate}

\section{DETAILED DESCRIPTION OF THE INVENTION}

\subsection{4.1~~Blind-Spot Decomposition Theory (BSDT)}

Given a trained classifier $f$ and a data instance $x$, the system computes
four independent diagnostic scores calibrated on known-normal data:

\begin{enumerate}
\item \textbf{Camouflage Score} $\delta_C(x)$: The Mahalanobis distance from
  the normal-period distribution:
  \[
    \delta_C(x) = \sqrt{(x - \mu_0)^\top \Sigma_0^{-1} (x - \mu_0)}
  \]
  where $\mu_0, \Sigma_0$ are the mean and (shrinkage-regularised) covariance
  of the normal-period data.

\item \textbf{Feature Gap Score} $\delta_G(x)$: The PCA residual norm in
  low-variance directions:
  \[
    \delta_G(x) = \|x - V_k V_k^\top x\|
  \]
  where $V_k$ are the top-$k$ eigenvectors of~$\Sigma_0$.

\item \textbf{Activity Anomaly Score} $\delta_A(x)$: The excess velocity
  beyond a percentile threshold:
  \[
    \delta_A(x) = \max\bigl(0, \, \|x_t - x_{t-1}\| - v_0\bigr)
  \]
  where $v_0 = \mathrm{percentile}_{95}(\|X_t - X_{t-1}\|)$ on the normal
  period.

\item \textbf{Temporal Novelty Score} $\delta_T(x)$: The kernel density
  self-history novelty score.
\end{enumerate}

The \textbf{Missed Fraud Likelihood Score} is the weighted composite:
\[
  \mathrm{MFLS}(x) = \sum_{i \in \{C,G,A,T\}} w_i \, \delta_i(x)
\]
with self-calibrating weights computed via one of: mutual information
weighting, Fisher variance ratio, or Bayesian online updating.

The \textbf{post-hoc correction formula} adjusts classifier outputs:
\[
  p^*(x) = p(x) + \lambda \cdot \mathrm{MFLS}(x) \cdot (1 - p(x))
  \cdot \mathbf{1}[p(x) < \tau]
\]
where $p(x)$ is the original classifier output, $\lambda$ is a scaling
parameter, and $\tau$ is a confidence threshold.

The \textbf{blind-spot energy functional} is:
\[
  E_{\mathrm{BS}}(X) = \delta_C^2 + \delta_G^2 + \delta_A^2 + \delta_T^2
\]

\subsection{4.2~~Universal Deviation Principle (UDP)}

The system employs a family of deviation-separating operators
$\{\Phi_k\}_{k=1}^K$ spanning five mathematical families:

\begin{enumerate}
\item \textbf{Statistical Spectrum:} Mahalanobis distance, z-score,
  interquartile deviation, copula rank
\item \textbf{Chaos/Dynamical Spectrum:} Lyapunov exponent, recurrence
  quantification, permutation entropy
\item \textbf{Spectral/Fourier Spectrum:} Power spectral density ratio,
  wavelet detail energy, Fourier phase coherence
\item \textbf{Geometric Spectrum:} Local intrinsic dimensionality, persistent
  homology, curvature
\item \textbf{Exponential Spectrum:} Tail index, extreme value deviation,
  L\'evy stability index
\end{enumerate}

The system guarantees detection via the \textbf{full-rank condition}: for the
stacked Jacobian $J = [D\Phi_1; \ldots; D\Phi_K]$, if $\mathrm{rank}(J) = d$
(the ambient dimension), then no anomaly can simultaneously match normal data
across all operators.

\textbf{Optimal fusion} weights $\{w_k\}$ are computed via semidefinite
programming:
\[
  \max_{w} \min_{\|v\|=1} \sum_k w_k \, v^\top D\Phi_k^\top D\Phi_k \, v
  \quad \text{subject to} \quad w_k \geq 0, \; \sum_k w_k = 1
\]

\textbf{Greedy operator selection} achieves a $(1 - 1/e)$-approximation via
submodular optimisation for selecting a size-$m$ subset of operators.

\subsection{4.3~~GravityEngine: N-Body Variational Anomaly Scoring}

Data points are modelled as particles evolving under three forces:

\begin{enumerate}
\item \textbf{Radial anchoring:} $F_{\mathrm{anchor}}(x_i) = -\alpha(x_i - \mu)$
  pulling each particle toward the data centroid.

\item \textbf{Pairwise interaction:} For each pair $(i,j)$, the interaction
  kernel combines attraction and repulsion:
  \[
    K(r) = -\exp\!\bigl(-r^2/\sigma^2\bigr) + \frac{\lambda_{\mathrm{rep}}}{r + \varepsilon}
  \]
  The attraction potential is:
  \[
    V_{\mathrm{attract}}(r) = \gamma \cdot \frac{\sigma\sqrt{\pi}}{2}
    \cdot \mathrm{erf}\!\bigl(r/\sigma\bigr)
  \]

\item \textbf{Deviation penalty:} $F_{\mathrm{dev}}(x_i)
  = -\sum_k \beta_k D\Phi_k(x_i)^\top D_k(x_i)$ coupling the particle
  dynamics to the UDP operator outputs.
\end{enumerate}

The system satisfies \textbf{force--energy consistency}: $F = -\nabla E$ for
a total energy $E = E_{\mathrm{anchor}} + E_{\mathrm{interact}} + E_{\mathrm{dev}}$.

Evolution proceeds via discrete integration with \textbf{backtracking
line-search} ensuring monotone energy descent at each step.  The anomaly score
for each point is its displacement $\|x_i^{(T)} - x_i^{(0)}\|$ after $T$
integration steps.

\subsection{4.4~~Boundary-Centred Dimension Magnifier}

For a representation matrix $R \in \mathbb{R}^{N \times D}$ with labels
$y \in \{0,1\}^N$, the transform processes each dimension~$d$ independently:

\begin{enumerate}
\item Compute boundary: $b_d = (\mu_{\mathrm{normal},d} + \mu_{\mathrm{anomaly},d})/2$
\item Standardise: $z = (r_d - b_d) / \sigma_{\mathrm{pooled},d}$
\item Compute steepness: $k_d = \gamma \cdot \mathrm{disc\_score}_d$
\item Apply saturating transform: $r'_d = \tanh(k_d \cdot z)$
\end{enumerate}

For discriminative dimensions (high $k_d$), classes saturate to $\pm 1$ with
amplified Cohen's~$d$.  For non-discriminative dimensions (low $k_d$), the
transform approaches zero, effectively silencing the dimension.

\subsection{4.5~~Risk-Aware Transaction Protocol (AMTTP)}

The protocol operates as a system of interconnected smart contracts on
EVM-compatible blockchains:

\begin{enumerate}
\item An \textbf{ML Oracle} computes off-chain risk scores in the range
  $[0, 1000]$ and signs them cryptographically (ECDSA).

\item The \textbf{Risk Router} validates the oracle signature and routes
  transactions to one of four tiers:
  \begin{itemize}
    \item Score $< 300$: \textbf{Approve} --- immediate execution
    \item Score $300$--$699$: \textbf{Review} --- enhanced monitoring
    \item Score $700$--$899$: \textbf{Escrow} --- funds locked in HTLC;
      multi-approver release required
    \item Score $\geq 900$: \textbf{Block} --- transaction rejected
  \end{itemize}

\item A \textbf{Circuit Breaker} pauses processing when $n$ consecutive
  transactions exceed the high-risk threshold.

\item A \textbf{Policy Engine} enforces velocity limits (daily/weekly volume
  caps), geographic restrictions, whitelist/blacklist management, and
  adaptive threshold adjustment.

\item \textbf{Hash Time-Locked Contracts (HTLC)} hold escrowed funds with
  cryptographic preimage release conditions and time-based expiry.
\end{enumerate}

\subsection{4.6~~Zero-Knowledge Non-Disclosing Anti-Fraud (zkNAF)}

The system implements privacy-preserving compliance using Groth16 ZK-SNARK
proofs on the BN254 elliptic curve.  Five proof types are supported:

\begin{enumerate}
\item \textbf{Sanctions Non-Membership:} Proves an address is not on a
  sanctions list via Merkle non-membership proof without revealing the address
  or list contents.

\item \textbf{Risk Range (Low/Medium):} Proves a transaction's ML risk
  score falls within an acceptable range without revealing the exact score.

\item \textbf{KYC Verified:} Proves KYC completion status without
  disclosing personal identity documents.

\item \textbf{Combined Transaction Compliance:} Composite proof combining
  all of the above.
\end{enumerate}

Verification requirements are tiered by transfer amount.  Proof records are
stored on-chain with expiry timestamps and revocation capability.

\subsection{4.7~~Cross-Chain Risk Propagation}

Risk scores, address block-lists, policy updates, and dispute results are
propagated across heterogeneous blockchains using a cross-chain messaging
protocol (LayerZero).  The system provides:

\begin{enumerate}
\item Per-chain rate limiting to prevent denial-of-service
\item Independent per-chain pause capability
\item Failed message retry with replay protection via incrementing nonces
\item Aggregation of risk scores from multiple source chains
\item Batch operations capped at 20 chains per call
\end{enumerate}

\subsection{4.8~~UI Integrity Verification Service}

A server-side verification system prevents UI manipulation attacks (such as
the February 2025 Bybit attack) by:

\begin{enumerate}
\item Computing cryptographic hashes of all page components (scripts, styles,
  forms, buttons) on the client side.

\item Verifying these hashes server-side against a trusted hash registry
  before executing any transaction.

\item Validating that the \textbf{transaction intent} (submitted data)
  matches the \textbf{displayed data hash} (what the user saw on screen).

\item Monitoring for real-time DOM mutations that may indicate injection
  attacks.

\item Per-component integrity checking with individual source, DOM, and
  event-handler hashes.
\end{enumerate}

\subsection{4.9~~Adaptive Friction Stability Framework}

The BSDT energy functional is applied to a dynamical system of networked
financial institutions with state vector $X$ encompassing leverage, correlation,
and network topology:

\begin{enumerate}
\item The system detects \textbf{spectral phase transitions} via the criterion:
  \[
    \lambda_{\max}\!\bigl(\rho(X) \, \ell(X) \, \mathbf{W}\bigr) = 1
  \]
  where $\rho$ is cross-correlation, $\ell$ is leverage, and $\mathbf{W}$ is
  the network adjacency matrix.

\item The \textbf{optimal adaptive damping coefficient} is derived in closed
  form:
  \[
    \gamma^*(X) \approx
    \frac{\beta \|\nabla E\|^2}{2\kappa + \beta \|\nabla E\|^2} \cdot
    \frac{E(X)}{E(X) + \theta}
  \]

\item An \textbf{online gradient descent} algorithm adjusts the coupling
  parameter in real time with $O(\sqrt{T})$ regret guarantee.

\item The system generates \textbf{countercyclical capital buffer (CCyB)
  recommendations} for macroprudential regulators.
\end{enumerate}

The equivalence between the BSDT blind-spot gradient and the Lyapunov
restoring force is established:
$\nabla E_{\mathrm{BS}} \equiv -F_{\mathrm{restore}}$.

\subsection{4.10~~State-Dependent Viscosity for Navier--Stokes Equations}

The BSDT energy functional is adapted for computational fluid dynamics.  Four
diagnostic channels are defined for velocity fields $\mathbf{u}$:

\begin{enumerate}
\item Enstrophy growth anomaly $\delta_C$
\item Spectral cascade deviation $\delta_G$
\item Vorticity--strain alignment anomaly $\delta_A$
\item Temporal novelty $\delta_T$
\end{enumerate}

The adaptive viscosity is:
\[
  \nu(\mathbf{u}) = \nu_0 \bigl(1 + \gamma^*(E_{\mathrm{BS}}(\mathbf{u}))\bigr), \qquad
  \gamma^*(E) = \frac{E}{E + \theta}
\]

This reduces the enstrophy production-to-dissipation ratio by approximately
50\% compared to constant viscosity at equivalent Reynolds numbers.

\subsection{4.11~~Compliance Orchestrator}

A central orchestration layer coordinates seven independent compliance
microservices in parallel:

\begin{enumerate}
\item ML Risk Scoring
\item Graph/Network Analysis
\item FCA Regulatory Compliance
\item Policy Management
\item Sanctions Screening
\item AML Pattern Monitoring (detecting structuring, layering, round-trip,
  fan-in/fan-out, peeling, and velocity anomalies)
\item Geographic Risk Assessment
\end{enumerate}

The orchestrator provides entity-profile-based routing (Retail, Institutional,
VASP, High-Net-Worth) where each profile type receives a different combination
of service checks.  Results are fused into a unified risk decision with
human-readable explainability output.

\newpage

%% ============================================================
%% CLAIMS
%% ============================================================
\section*{CLAIMS}

\bigskip
\textbf{I/We claim:}

\bigskip

%% --- GROUP A: BSDT ---
\begin{enumerate}[label=\textbf{\arabic*.}, leftmargin=2em]

\item A computer-implemented method for identifying missed detections of a
trained classification model, comprising:
\begin{enumerate}[label=(\alph*)]
  \item computing, for each data instance, four independent diagnostic scores
    calibrated solely on known-normal data: a camouflage score measuring
    Mahalanobis distance from normal-period distribution, a feature gap score
    measuring PCA residual norm in low-variance directions, an activity anomaly
    score measuring excess velocity beyond a percentile threshold, and a
    temporal novelty score measuring kernel-density self-history novelty;
  \item computing a composite missed-prediction likelihood score as a weighted
    sum of said four scores using self-calibrating weights; and
  \item adjusting the output probability of said trained classification model
    using said composite score according to the formula
    $p^*(x) = p(x) + \lambda \cdot \mathrm{MFLS}(x) \cdot (1-p(x)) \cdot
    \mathbf{1}[p(x) < \tau]$
    without retraining said model.
\end{enumerate}

\item The method of claim~1, wherein the self-calibrating weights are computed
using one of: mutual information weighting, Fisher variance ratio weighting, or
Bayesian online updating.

\item The method of claim~1, further comprising computing a blind-spot energy
functional $E_{\mathrm{BS}} = \delta_C^2 + \delta_G^2 + \delta_A^2 + \delta_T^2$
as a unified diagnostic signal.

\item A computer-implemented method for generating supplementary features for
a machine learning model from the four diagnostic scores of claim~1, comprising
deriving at least seven additional features from the camouflage, feature gap,
activity anomaly, and temporal novelty components and their interactions.

%% --- GROUP B: UDP ---
\item A computer-implemented method for anomaly detection, comprising:
\begin{enumerate}[label=(\alph*)]
  \item projecting data through a family of $K$ deviation-separating operators
    $\{\Phi_k\}_{k=1}^K$ spanning at least three of: statistical, dynamical,
    spectral, geometric, and exponential mathematical families;
  \item verifying that the stacked Jacobian
    $J = [D\Phi_1; \ldots; D\Phi_K]$ has rank equal to the ambient dimension,
    thereby guaranteeing that no anomaly can simultaneously match normal data
    across all operators; and
  \item computing a composite anomaly score as a weighted combination of
    operator outputs.
\end{enumerate}

\item The method of claim~5, wherein the weights are computed via semidefinite
programming maximising worst-case directional sensitivity subject to
non-negativity and unit-sum constraints.

\item The method of claim~5, further comprising selecting a size-$m$ subset of
operators using a greedy algorithm with a $(1-1/e)$-approximation guarantee
via submodular optimisation.

%% --- GROUP C: GravityEngine ---
\item A computer-implemented method for anomaly detection using particle
simulation, comprising:
\begin{enumerate}[label=(\alph*)]
  \item modelling each data point as a particle in a force field comprising
    radial anchoring toward a data centroid, pairwise interaction via kernels
    having attraction and repulsion components, and deviation penalty forces
    coupled to anomaly operator outputs;
  \item evolving particle positions via iterative integration with
    backtracking line-search ensuring monotone energy descent; and
  \item computing an anomaly score for each data point as its displacement
    from initial to final position after evolution.
\end{enumerate}

\item The method of claim~8, wherein the pairwise interaction kernel is
$K(r) = -\exp(-r^2/\sigma^2) + \lambda_{\mathrm{rep}}/(r + \varepsilon)$ and
the attraction potential is
$V_{\mathrm{attract}}(r) = \gamma \cdot (\sigma\sqrt{\pi}/2) \cdot
\mathrm{erf}(r/\sigma)$.

\item The method of claim~8, satisfying force--energy consistency
$F = -\nabla E$ for a total energy functional
$E = E_{\mathrm{anchor}} + E_{\mathrm{interact}} + E_{\mathrm{dev}}$.

%% --- GROUP D: Dimension Magnifier ---
\item A computer-implemented method for nonlinear feature transformation,
comprising, for each dimension $d$ of a feature representation:
\begin{enumerate}[label=(\alph*)]
  \item computing a decision boundary $b_d$ as the midpoint between mean
    normal and mean anomaly values;
  \item standardising each value by the pooled standard deviation;
  \item computing a steepness $k_d$ proportional to a per-dimension
    discriminative score; and
  \item applying a saturating transform $r'_d = \tanh(k_d \cdot z)$,
    thereby amplifying discriminative dimensions and silencing
    non-discriminative ones.
\end{enumerate}

%% --- GROUP E: AMTTP Protocol ---
\item A blockchain-based transaction processing system, comprising:
\begin{enumerate}[label=(\alph*)]
  \item an off-chain machine-learning oracle that computes a risk score for
    each transaction and signs the score cryptographically;
  \item an on-chain risk router that verifies the oracle signature and routes
    transactions to one of at least four risk tiers, wherein the highest tier
    blocks the transaction and an intermediate tier escrows funds in a hash
    time-locked contract requiring multi-party approval for release;
  \item a circuit breaker that automatically pauses transaction processing
    when a configurable number of consecutive transactions exceed a risk
    threshold; and
  \item a policy engine enforcing velocity limits, geographic restrictions,
    and adaptive threshold adjustment.
\end{enumerate}

\item The system of claim~12, further comprising a hash time-locked escrow
contract wherein funds are released upon presentation of a cryptographic
preimage or automatically refunded upon time-based expiry.

%% --- GROUP F: zkNAF ---
\item A computer-implemented system for privacy-preserving compliance
verification on a blockchain, comprising:
\begin{enumerate}[label=(\alph*)]
  \item generating zero-knowledge proofs using a Groth16 ZK-SNARK protocol
    on a BN254 elliptic curve for at least: sanctions list non-membership via
    Merkle non-membership proof, risk score range attestation, and KYC
    completion verification;
  \item verifying said proofs on-chain without the verifier learning the
    underlying personal data, sanctions list contents, or exact risk score;
  \item applying tiered verification requirements wherein the transfer amount
    determines which proofs are required; and
  \item storing proof records on-chain with expiry timestamps and revocation
    capability.
\end{enumerate}

%% --- GROUP G: Cross-Chain ---
\item A system for propagating machine-learning-derived risk scores across
multiple heterogeneous blockchains, comprising:
\begin{enumerate}[label=(\alph*)]
  \item encoding risk scores, address block/unblock commands, policy updates,
    and dispute results as typed cross-chain messages;
  \item transmitting said messages via a cross-chain messaging protocol with
    per-chain rate limiting and replay protection via incrementing nonces;
  \item aggregating risk scores received from multiple source chains for a
    single address; and
  \item providing independent per-chain pause capability.
\end{enumerate}

%% --- GROUP H: UI Integrity ---
\item A computer-implemented method for preventing user-interface manipulation
attacks on financial applications, comprising:
\begin{enumerate}[label=(\alph*)]
  \item computing, on a client device, cryptographic hashes of page components
    including scripts, stylesheets, forms, and interactive elements;
  \item transmitting said hashes to a server which verifies them against a
    trusted hash registry;
  \item validating that a transaction intent submitted by the user matches a
    displayed-data hash corresponding to the information shown on the user's
    screen, thereby ensuring the submitted transaction matches the displayed
    transaction; and
  \item monitoring for real-time DOM mutations indicating potential injection
    attacks.
\end{enumerate}

\item The method of claim~16, further comprising per-component integrity
checking with individual source hash, DOM hash, and event-handler hash for
each interactive element.

%% --- GROUP I: Adaptive Friction ---
\item A computer-implemented method for detecting systemic financial risk and
generating macroprudential policy recommendations, comprising:
\begin{enumerate}[label=(\alph*)]
  \item computing the blind-spot energy functional
    $E_{\mathrm{BS}} = \delta_C^2 + \delta_G^2 + \delta_A^2 + \delta_T^2$
    from financial institution state data encompassing leverage, correlation,
    and network topology, where the four diagnostic channels are calibrated
    solely on normal-market-period data;
  \item detecting spectral phase transitions via the criterion
    $\lambda_{\max}(\rho(X) \, \ell(X) \, \mathbf{W}) = 1$ where $\rho$ is
    cross-correlation, $\ell$ is leverage, and $\mathbf{W}$ is the network
    adjacency matrix;
  \item computing an optimal adaptive damping coefficient
    $\gamma^*(X) = \frac{\beta \|\nabla E\|^2}{2\kappa + \beta \|\nabla E\|^2}
    \cdot \frac{E(X)}{E(X) + \theta}$; and
  \item generating countercyclical capital buffer recommendations based on
    said damping coefficient.
\end{enumerate}

\item The method of claim~18, further comprising an online gradient descent
algorithm that adjusts the coupling parameter in real time with
$O(\sqrt{T})$ regret guarantee.

\item The method of claim~18, wherein the equivalence between the blind-spot
gradient and a Lyapunov restoring force is exploited:
$\nabla E_{\mathrm{BS}} \equiv -F_{\mathrm{restore}}$.

%% --- GROUP J: Navier-Stokes ---
\item A computer-implemented method for adaptive viscosity in computational
fluid dynamics, comprising:
\begin{enumerate}[label=(\alph*)]
  \item computing, from a velocity field $\mathbf{u}$, four diagnostic channel
    scores: enstrophy growth anomaly, spectral cascade deviation,
    vorticity--strain alignment anomaly, and temporal novelty;
  \item computing a composite blind-spot energy
    $E_{\mathrm{BS}}(\mathbf{u}) = \delta_C^2 + \delta_G^2 + \delta_A^2
    + \delta_T^2$; and
  \item applying a state-dependent viscosity
    $\nu(\mathbf{u}) = \nu_0(1 + \gamma^*(E_{\mathrm{BS}}(\mathbf{u})))$
    where $\gamma^*(E) = E/(E + \theta)$, thereby reducing enstrophy
    production-to-dissipation ratio without requiring manual tuning.
\end{enumerate}

%% --- GROUP K: Orchestrator ---
\item A computer-implemented compliance system comprising:
\begin{enumerate}[label=(\alph*)]
  \item a central orchestration layer that invokes, in parallel, at least five
    independent compliance microservices including machine-learning risk
    scoring, graph analysis, sanctions screening, AML pattern monitoring, and
    geographic risk assessment;
  \item entity-profile-based routing wherein each entity type receives a
    different combination of compliance checks; and
  \item a fusion module that combines the outputs of said microservices into
    a unified risk decision with human-readable explainability.
\end{enumerate}

\item The system of claim~22, wherein the AML pattern monitoring service
detects at least: structuring/smurfing, layering, round-trip transactions,
fan-in/fan-out patterns, peeling chains, and velocity anomalies using
adaptive thresholds.

%% --- GROUP L: Cross-Domain Unification ---
\item A computer-implemented method for cross-domain anomaly detection and
stability control using a unified blind-spot energy functional, wherein the
same functional $E_{\mathrm{BS}} = \sum_{i \in \{C,G,A,T\}} \delta_i^2$ is
applied, with domain-specific channel definitions, to at least two of:
financial fraud detection, macroeconomic systemic risk monitoring, blockchain
transaction compliance, and computational fluid dynamics viscosity control.

%% --- GROUP M: Dispute Resolution ---
\item A system for resolving disputes on machine-learning-flagged transactions
using decentralised arbitration, comprising:
\begin{enumerate}[label=(\alph*)]
  \item automatically escrowing funds when a machine-learning model assigns a
    risk score above a configurable threshold;
  \item providing a challenge window during which either party may initiate a
    dispute;
  \item submitting structured ML evidence to a decentralised arbitration
    tribunal; and
  \item executing the tribunal's ruling on-chain, supporting partial
    percentage-based splits.
\end{enumerate}

\end{enumerate}

\newpage

%% ============================================================
%% ABSTRACT (repeated for Form 2)
%% ============================================================
\section*{ABSTRACT OF THE DISCLOSURE}

A unified system and method for detecting anomalies, scoring risk, enforcing
compliance, and controlling stability across heterogeneous domains including
financial fraud detection, blockchain transaction processing, macroeconomic
stability, and computational fluid dynamics.  The system is built upon a
Blind-Spot Decomposition Theory (BSDT) that decomposes classification failures
into four independent diagnostic channels---Camouflage, Feature Gap, Activity
Anomaly, and Temporal Novelty---producing a composite energy functional.  This
functional drives: a Universal Deviation Principle for multi-spectrum anomaly
detection with provable guarantees; an N-body variational scoring engine with
closed-form energy potentials; a blockchain transaction protocol with tiered
risk-aware escrow and zero-knowledge compliance proofs; a Lyapunov-based
adaptive friction framework for systemic risk detection with closed-form
optimal macroprudential policy; and a state-dependent viscosity method for
Navier--Stokes simulation.\quad (148 words)

\end{document}
